---
title: "Data Analysis Project"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE,
                      comment = "")
```



# Student Information

**Name**: Deepthi Tamma
**G Number**: G01241465




# Introduction

This semester we will be working with a data set from the field of **Human Resources Analytics**.

Broadly speaking, this field is concerned with using employee data within a company to optimize objectives such as employee satisfaction, productivity, project management, and most commonly, avoiding employee attrition. 

Ideally, companies would like to keep **attrition rates** (the proportion of employees leaving a company for other opportunities) as low as possible due to the variable costs and business disruptions that come with having to replace productive employees on short notice.

The objective of this project is to perform an exploratory data analysis on the `employee_data` data set to uncover potential solutions for minimizing employee attrition rates. 



## Employee Attrition Data

The `employee_data` data frame is loaded below and consists of 1,470 employee records for a U.S. based product company. The rows in this data frame represent the attributes of an employee at this company across the variables listed in the table below.

**Note**: If you have not installed the `tidyverse` package, please do so by going to the `Packages` tab in the lower right section of RStudio, select the `Install` button and type `tidyverse` into the prompt.


```{r}
library(tidyverse)

employee_data <- read_rds(url('https://gmubusinessanalytics.netlify.app/data/employee_data.rds'))
```


### Raw Data

```{r}

employee_data

```
<br>
<br>


# Exploratory Data Analysis


Executives at this company have hired you as a data science consultant to identify the factors that lead to employees leaving their company. 

They would like for you to explore why employees are leaving their company and make recommendations on how to minimize this behavior.

You must think of at least 8 relevant questions that explore the relationship between `left_company` and the other variables in the `employee_data` data frame. 

The goal of your analysis should be discovering which variables drive the differences between employees who do and do not leave the company.

You must answer each question and provide supporting data summaries with either a summary data frame (using `dplyr`/`tidyr`) or a plot (using `ggplot`) or both.

In total, you must have a minimum of 5 plots and 4 summary data frames for the exploratory data analysis section. Among the plots you produce, you must have at least 4 different types (ex. box plot, bar chart, histogram, heat map, etc...)


Each question must be answered with supporting evidence from your tables and plots. See the example question below.


## Sample Question

Is there a relationship between employees leaving the company and their current salary?



**Answer**: Yes, the data indicates that employees who leave the company tend to have lower salaries when compared to employees who do not. Among the 237 employees that left the company, the average salary was \$76,625. This is over \$20,000 less than the average salary of employees who did not leave the company.

Among the employees *who did not leave the company*, only 10% have a salary that is less than or equal to \$60,000. When looking at employees who did leave the company, this increase to 34%.


### Summary Table

```{r}
employee_data %>% group_by(left_company) %>% 
                  summarise(n_employees = n(),
                            min_salary = min(salary),
                            avg_salary = mean(salary),
                            max_salary = max(salary),
                            sd_salary = sd(salary),
                            pct_less_60k = mean(salary <= 60000))
```


### Data Visulatization

```{r}
ggplot(data = employee_data, aes(x = salary, fill = left_company)) + 
   geom_histogram(aes(y = ..density..), color = "white", bins = 20) +
   facet_wrap(~ left_company, nrow = 2) +
   labs(title = "Employee Salary Distribution by Status (Left the Comapny - Yes/No)",
           x = "Salary (US Dollars", y = "Proportion of Employees")
```




To add additional R code chunks for your work, select `Insert` then `R` from the top of this notebook file.


## Question 1



**Question**: Is there any relation between the employees leaving the company and the number of working hours?



**Answer**: Yes, the plot shows that the employees tend to leave the company when the number of working hours are increased showing a negative impact of increased stress and work pressure. 




```{r}
ggplot(data=employee_data, aes(x=weekly_hours))+geom_histogram(bins=20)+
  facet_wrap(~ left_company)+scale_y_log10() + labs(title = "Distribution of Weekly hours by Number of employees",x= "Weekly hours", y="No. of Employees")


```



## Question 2



**Question**: Is there any relationship between frequency of the employees leaving their previous companies and them leaving the current ocmpany? 



**Answer**: From the box plot, it can be infered that employees who tend to have more number of previously worked companies tend to change companies often which shows that the probabilitity of the employee leaving the current company is high.



```{r}
ggplot(employee_data,aes(x=left_company,y=previous_companies))+geom_boxplot(fill="#006EA1")+
  scale_y_log10()+ labs(title = "Box plot of Left company by Previous Companies",
                        x= "Left company",y="No of Previous Companies worked")



```




## Question 3



**Question**: Is there any relationship between employee leaving company with respect to the salary and number of years at the company? 



**Answer**: Yes, it can be infered from the plot that employees who do not get a raise in the salaries over the years of their service have higher probability of leaving the company.




```{r}
ggplot(employee_data, aes(x=yrs_at_company, y=salary,color=left_company)) + 
  geom_point()+facet_grid(~ left_company)+labs(title = "No. of years at company vs Salary",
                                               x= "No. of years at company",
                                               y="Salary")

```




## Question 4



**Question**: Is there any relationship between performance rating and number of years worked with respect to job satisfaction?



**Answer**: From the plot, it can be understood that though the employees might have exceptional performance rating, they might not have higher job satisfaction rates which might lead them to leave the company as the years progress. 


```{r}
by_rating <- employee_data %>% group_by(left_company, performance_rating, job_satisfaction) %>%
summarize(no_of_yrs = mean(yrs_at_company))
```

```{r} 
ggplot(by_rating,aes(x = performance_rating, y = no_of_yrs )) + 
  geom_col(aes(fill = job_satisfaction),width = 0.7) + facet_wrap(~left_company)



```

## Question 5



**Question**: Is there any relationship between employees leaving the company with Number of Years at company and number of years since promotion?



**Answer**: yes, itseems that the employees who worked for longer time didn't get any promotion when compared with those who did not leave the company. 



```{r}
ggplot(employee_data, aes(x=yrs_at_company, y=yrs_since_promotion)) + 
  geom_point(color="#006EA1",size=2,alpha=0.50)+facet_wrap(~ left_company)+
labs(title = "Years at Company Vs Years since Promotion ",
     x="Years at Company",
     y="Years since Promotion")


```

## Question 6



**Question**: Is there any relationship between employees leaving the company due to Performance rating and salary ?



**Answer**: Yes, even though the ratings are same for the employees some of the employees are given more salary and this makes them leave the comapny. 

```{r}
by_rating <- employee_data %>% group_by(left_company,performance_rating) %>% 
  summarise(avg_salary = mean(salary))
by_rating

```

```{r}
ggplot(by_rating,aes(x=performance_rating,y=avg_salary)) +
  geom_col() + facet_wrap(~ left_company) + labs(title = "Performance Rating vs Salary",
                                                 x="Performance Rating",
                                                 y="Salary")


```

## Question 7



**Question**: Is there any relation between employees leaving the company due to salary depending on the department?



**Answer**: Yes, it can be observed that the employees who left the company working under various departments have less average salaries when compared to those who didnot leave the company. For example, in the Marketting and Research department, the difference in average salary of the employees who left the company is over $30,000




```{r}
employee_data%>%group_by(left_company,department)%>%
                summarise(n_employees = n(),
                          avg_salary=mean(salary),
                          max_salary=max(salary),
                          sd_salary=sd(salary))


```



## Question 8



**Question**: Is there any relationship between the employees leaving company due to the distance to be travelled?



**Answer**: Yes, the data indicates that the employees tend to leave the company when the distance to be travelled is more. It can be observed that the average distance is less for employees who are still working at the company.




```{r}
employee_data %>% group_by(left_company) %>% 
                  summarise(n_employees = n(),
                          min_distance = min(miles_from_home),
                          avg_distance = mean(miles_from_home),
                          max_distance = max(miles_from_home),
                          sd_distance = sd(miles_from_home)
                         )


```
## Question 9



**Question**: Is there any relationship between the weekly working hours and salary with the employee leaving the company?



**Answer**: Yes, it can be infered that the employees who left the company have worked for more number of hours for less salary. This is over $20,000 less than the average salary of employees who did not leave the company.




```{r}
employee_data %>% group_by(left_company) %>%
  summarise(n_employees = n(),
            avg_hrs=mean(weekly_hours),
            max_hrs=max(weekly_hours),
            avg_salary=mean(salary),
            max_salary = max(salary),
            )

```

## Question 10



**Question**: Is there any relationship between employees leaving the company with respect to marital status and distance between the workplace and home?



**Answer**: No, this doesn't much affect the employee decision to change the company since the distance is pretty much same for the employees who still work and those who left the company.



```{r}
employee_data %>% group_by(left_company,marital_status) %>%
  summarise(n_employees = n(),
            avg_distance=mean(miles_from_home))

```

## Question 11



**Question**: Is there any relationship between employees leaving the company with respect to Job Level status and No. of previous companies they worked for?



**Answer**: yes, the data shows that the employees who left the current company at different job levels tend to change their companies often. It can be seen that all those employees who left the company worked for atleast 4 previous companies.

```{r}
employee_data %>% group_by(left_company,job_level)%>%
  summarise(n_employees = n(),
            avg_no_of_companies=mean(previous_companies))

```

# Summary of Results

Write an executive summary of your overall findings and recommendations to the executives at this company. Think of this section as your closing remarks of a presentation, where you summarize your key findings and make recommendations to improve HR processes at the company. 

Your executive summary must be written in a [professional tone](https://www.universalclass.com/articles/writing/business-writing/appropriate-tone-in-business-communications.htm), with minimal grammatical errors, and should include the following sections:

1. An introduction where you explain the business problem and goals of your data analysis

    - What problem(s) is this company trying to solve? Why are they important to their future success?
  
    - What was the goal of your analysis? What questions were you trying to answer and why do they matter?


2. Highlights and key findings from your Exploratory Data Analysis section 
    - What were the interesting findings from your analysis and **why are they important for the business**?

    - This section is meant to **establish the need for your recommendations** in the following section


3. Your recommendations to the company on how to reduce employee attrition rates 
  
    - Each recommendation must be supported by your data analysis results 

    - You must clearly explain why you are making each recommendation and which results from your data analysis support this recommendation

    - You must also describe the potential business impact of your recommendation:
      
      - Why is this a good recommendation? 
      
      - What benefits will the business achieve?
      


## Executive Summary

Please write your executive summary below. If you prefer, you can type your summary in a text editor, such as Microsoft Word, and paste your final text here.

Retention of productive employees is a significant concern of business executives for a company to be fruitful. There are many key factors which affect the productivity and one of those is the control of employee attrition rates. The main goal of the analysis is to analyze the factors which affect the employee attrition and take the respective measures to reduce them. Here, we are analyzing on the factors such as Job Satisfaction, Performance Rating, Salary, Job level, marital status, number of weekly hours, frequency of business travel, number of years at the company, number of years since promotion, number of previous companies the employee worked for and the distance from home.

Employees who leave the organization varies for different reasons and from the data analysis, it is found that the most important factors that affect are Salary, number of working hours, Job satisfaction, performance rating, distance from home. So, the company should focus on these factors to minimize the attrition rates. 

The foremost reason the employees tend to leave the company is due to the number of weekly hours. From the analysis, it can be said that most of the employees are working for longer duration due to which there is increased work pressure. 

Another factor which has a greater impact on employee attrition is Salary. Though the employees have been in the company for a longer period of time, there were no hikes in the salary for which they left the organization. 

Distance between the workplace to home also plays a major role in attrition rates. It can be seen from the analysis that the employees whose home is farther from the workplace left the company. 

Job Satisfaction along with the performance rating is also another key factor that is affecting the employee attrition. Some of employees left the organization even though their performance was exceptional because they had less job satisfaction. In another case, employees’ performance was not effective for which they left the organization.



Some of the interesting findings of my analysis is with Business travel which is not a considerable factor for an employee to leave the company. There are many employees’ who didn’t leave the organization even though there was rarely to no business travel. 

Number of previously worked companies shows that those employees who have frequently changed the companies will tend to leave the organization. 

Performance rating and Salary affect the employee attrition rates very much. In the analysis, it is found that though the ratings are sane for employees there is bias in the salary distribution which would be discouraging and that leads to attrition.

It is also observed that years since promotion affects the employee attrition rate. Employees who were not promoted even if they worked for longer duration at the company left the organization. 

There is also difference in the salary of the employees who work for the various departments. The difference in price is over $30,000 between who work for the same department which is discouraging and shows partiality towards employees and hence left the organization.

The distance between the workplace and marital status doesn’t much affect the decision of an employee to leave the company since the difference in the distance is low when compared to the employees who work for the organization irrespective of the marital status. 

Employee retention involves taking measures to encourage employees to remain in the organization for the maximum period of time. 

One of the most important recommendations to reduce employee attrition rates is to provide salary hikes to deserving employees.
Benefit: This helps the employees to stay motivated and work with better enthusiasm.

Another main recommendation can be to reduce the weekly working hours of the employees.
Benefit: It has been proven scientifically that a person's concentration level is the highest when they are stress-free i.e., reducing the working hours can help obtain a healthy, stress-free work environment which in turn improves the effectiveness of the employees. This, in the long run, aids in the company's profits.

Lastly, having bias-free and appropriate salaries for every department and depending upon the ratings is recommended.
Benefit: Having such a bias-free environment, for salaries as well as other employee-benefits assists in inter-departmental synergy and helps the employees work together as a team for the company.
































